---
title: "Homicide Trends"
author: "290353"
format: revealjs
---

## Introduction

This slide deck summarizes key insights from homicide data across four European cities between 2000 and 2007.

---

## Dataset Overview

- **Source**: Eurostat (`estat_crim_hom_city_filtered_en.csv.gz`)
- **Cities**: Berlin, Warsaw, Ankara, London
- **Years**: 2000 to 2007

---

## Trend Over Time

```{r}
library(dplyr)
library(ggplot2)

data <- read.csv("estat_crim_hom_city_filtered_en.csv.gz")

clean_data <- data %>%
  select(city = Geopolitical.entity..declaring., year = TIME_PERIOD, homicides = OBS_VALUE)

ggplot(clean_data, aes(x = year, y = homicides, color = city)) +
  geom_line(size = 1.2) +
  geom_point() +
  labs(
    title = "Homicide Trends (2000–2007)",
    x = "Year",
    y = "Number of Homicides",
    color = "City"
  ) +
  theme_minimal()
```

---

## Average Homicides per City

```{r}
avg_data <- clean_data %>%
  group_by(city) %>%
  summarise(avg_homicides = mean(homicides, na.rm = TRUE))

ggplot(avg_data, aes(x = city, y = avg_homicides, fill = city)) +
  geom_col() +
  labs(
    title = "Average Homicides per City (2000–2007)",
    x = "City",
    y = "Average Number of Homicides"
  ) +
  theme_minimal()
```

---

## Total Homicides by Year

```{r}
total_by_year <- clean_data %>%
  group_by(year) %>%
  summarise(total = sum(homicides, na.rm = TRUE))

ggplot(total_by_year, aes(x = year, y = total)) +
  geom_line(color = "steelblue", size = 1.2) +
  geom_point(color = "steelblue") +
  labs(
    title = "Total Homicides Across All Cities (2000–2007)",
    x = "Year",
    y = "Total Number of Homicides"
  ) +
  theme_minimal()
```

---

## Conclusion

- Berlin showed a significant decline in homicides  
- Ankara had consistently high numbers  
- Overall homicide levels in selected cities decreased

Thank you!
